Homework

### 2016-06-29 Macro Impacts.tex
This had better work
OR ELSE

% Title
\title{Macro Impacts}
\author{David Cyncynates \\ dcc57@case.edu}
\date{\today}

\begin{document}
\maketitle
The linearized elastic wave equations are (Hemmann)
\begin{align}
0&=(\lambda+2\mu)\nabla^2\phi-\rho\ddot\phi+f\,,\\
0&=\mu\nabla(\nabla\cdot A)-\mu\nabla\times(\nabla\times A)-\rho\ddot A+G\,,
\end{align}
where $u=\nabla\phi+\nabla\times A$ is the displacement and $F=\nabla f+\nabla\times G$ is the body force. Hence, we see to first order that shear waves and compressional waves decouple in the linear regime. If we assume that displacements due to the impact are very small (as we would expect for a very dense macro impact), then we may assume hydrostatic equilibrium, in which case the stress tensor takes the form
\begin{align}
-p\delta_{ij}&=\tau_{ij}=\lambda g_{ij} u^s_{,s}+\mu(u_{i,j}+u_{j,i})\,.
\end{align}
Note, the commas denote covariant differentiation. We now transform into cylindrical coordinates. In the case that we are interested in, the only important component is the radial one, for which we have
\begin{align}
-p&=\tau_{rr}\,,\\
&=\lambda u^s_{,s}+2\mu u_{r,r}\,,\\
&=\lambda \nabla_{s}g^{ss}u_s+2\mu u_{r,r}\,.
\end{align}
The metric of cylindrical coordinates is
\begin{align}
\diff s^2&=\diff r^2+r^2\diff\theta^2+\diff z^2\,,
\end{align}
leading to the Christoffel symbols
\begin{align}
\Gamma^r_{\theta\theta}&=-r\,,\\
\Gamma^r_{r\theta}&=\f1r\,.
\end{align}
Thus, assuming only the radial component has non-zero derivatives (which will be the case for a cylindrically symmetric source),
\begin{align}
\nabla u&=\partial_r u_r+\f1r u_r\,,\\
&=\f1r\partial_r(ru_r)\,.
\end{align}
Hence
\begin{align}
-p&=\tau_{rr}=\lambda\f1r\partial_r(ru_r)+2\mu\partial_r u_r\,,\\
&=\lambda\f1r\partial_r(ru_r)+2\mu\f1r(\partial_r ru_r)-2\mu\f1r(u_r)\,,\\
&=(\lambda+2\mu)\f1r\partial_r(ru_r)-2\mu\f1ru_r\,,\\
&=(\lambda+2\mu)\f1r\partial_r(r\partial_r\phi)-2\mu\f1r\partial_r\phi
\end{align}
With $f=0$, this and the scalar equation of motion yield
\begin{align}
0&=(\lambda+2\mu)\nabla^2\phi-\rho\ddot\phi\,,\\
&=(\lambda+2\mu)\p{\f1r\partial_r\p{r\partial_r\phi}}-\rho\ddot\phi\,,\\
&=-p+2\mu\f1r\partial_r\phi-\rho\ddot\phi\,,
\end{align}
Thus, observe that all the solutions from Hemmann apply with small modification (compare 19 with his equation 9).
Assuming only outgoing modes are generated, we may write $\phi=F(t-r/\alpha)$ where $\alpha=\sqrt{(\lambda+2\mu)/\rho}$ is the compressional wave velocity. We can, for the sake of reusing Hemmann's work, assume that the cavity produced by the macro existed prior to the generation of compressional modes. This is actually somewhat convenient, since it is the production of the cavity that generates shear modes. Denote differentiation with respect to the retarded time $t'=t-\f r\alpha$ by a prime. I can't figure out how to get the following from equation 11 in Hemmann, but the evolution equation is
\begin{align}
-r^3\alpha p&=\alpha\rho r^2 F''+2\mu r F'+2\mu\alpha F\,.
\end{align}
After some guessing and checking, it seems that one can arrive at this equation by the substitution $\phi=\f1rF(t-r/\alpha)$. Either way, this simply confirms that we can use Hemmann directly with only slight modification.
\end{document}

>>>>

% Title
\title{Homework}
\author{David Cyncynates \\ dcc57@case.edu}
\date{\today}

\begin{document}
\maketitle
In the following, lower case denotes a quantity in position space while capital letters denote their components in Fourier space.
\\\\
Denote the displacement field $u(\vec x,t)=\nabla\phi(\vec x,t)+\nabla\times a(\vec x,t)$\,. The linearized (acoustic) wave equation for $\phi$ is then
\begin{align}
\alpha^2\nabla^2\phi&=\partial_t^2\phi\,,
\end{align}
where $\alpha^2=\f{\lambda+2\mu}{\rho}$. We may write
\begin{align}
\phi(\vec x,t)=\int\f{\diff^3 k}{(2\pi)^3}\Phi(\vec k,\omega)e^{\I(\vec k\cdot\vec x-\omega t)}\,,
\end{align}
from which we obtain the dispersion relation $\alpha k=\omega$, and hence the solution
\begin{align}
\phi(\vec x,t)=\int\f{\diff^3 k}{(2\pi)^3}\Phi(\vec k,\omega)e^{\I(\vec k\cdot\vec x-\alpha k t)}\,.
\end{align}
We now impose the constraint equation
\begin{align}
\sigma_{ij}&=\delta_{ij}\lambda\nabla\cdot u+\mu(u_{i,j}+u_{j,i})
\end{align}
which, for the scalar modes becomes
\begin{align}
\sigma_{ij}&=\delta_{ij}\lambda\nabla^2\phi+2\mu\partial_i\partial_j\phi
\end{align}
and whose trace is
\begin{align}
-p=K\nabla^2\phi\,,
\end{align}
where $K=\lambda+\f23\mu$ is the bulk modulus and $p=-\f13\tr\sigma_{ij}$. We take this constraint as an initial condition at $t=0$. In Fourier space
\begin{align}
P(\vec k)=Kk^2\Phi(\vec K)\,,
\end{align}
from which we obtain the displacement field components
\begin{align}
U(\vec k)&=\I\f{P(\vec k)}{K}\f{\vec k}{k^2}\,,
\end{align}
and hence
\begin{align}
u(\vec x,t)=\int\f{\diff^3 k}{(2\pi)^3}\I\f{P(\vec k)}{K}\f{\vec k}{k^2}e^{\I(\vec k\cdot\vec x-\alpha k t)}\,,
\end{align}
The equation for the energy of the compressional modes is
\begin{align}
E&=\f12\int_V\diff^3x\p{\rho\abs{\partial_t u}^2+(\lambda+2\mu)\abs{\nabla\cdot u}^2}\,,\\
&=\f{\lambda+2\mu}{K^2}\int\f{\diff^3 k}{(2\pi)^3}\abs{P(\vec k)}^2
\end{align}
In the case that $P$ depends only on the frequency
\begin{align}
E&=4\pi\f{\lambda+2\mu}{K^2}\int\f{\diff\tilde\lambda}{\tilde\lambda^4}\abs{P(\tilde\lambda)}^2\,.
\end{align}
Now consider the case of a step function type pressure source - a cylinder of height $h$ and radius $r_X$
\begin{align}
p(\vec x,0)&=p_0\theta(r_X-r)\ps{\theta(z+h/2)-\theta(z-h/2)}\,,
\end{align}
whose Fourier components are
\begin{align}
P(\vec k)&=\f{4\pi r_X p_0}{\sqrt{k_x^2+k_y^2}k_z}J_1\p{\sqrt{k_x^2+k_y^2}r_X}\sin\p{\f{h}{2}k_z}\,,
\end{align}
which, in polar $k$-space is
\begin{align}
P(k,\varphi,\theta)&=\f{4\pi r_X p_0}{k^2\sin\theta\cos\theta}J_1\p{r_Xk\sin\theta}\sin\p{\f{h}{2}k\cos\theta}\,,
\end{align}
Directly integrating in Mathematica (over $\theta$ first) yields
\begin{align}
E_{\te{total}}&=\f{\lambda+2\mu}{K^2}4\pi^2p_0^2\sigma_Xh
\end{align}
where $\sigma_X=\pi r_X^2$. To calculate the energy deposition into the low frequency spectrum, we integrate $k$ from $0$ to $k_0$. Observe, we can make the following approximation
\begin{align}
P(\vec k)&\approx_{k\ll r_X}\f{2\pi r_Xp_0}{k^2\cos\theta\sin\theta}r_X k\sin\theta\sin\p{\f{h}{2}k\cos\theta}\,,\\
&=\f{2\pi r_X^2 p_0}{k\cos\theta}\sin\p{\f{h}{2}k\cos\theta}\,.
\end{align}
From this we obtain the portion of the energy relegated to the long wavelength spectrum
\begin{align}
E_{\te{propagated}}&=\f{\lambda+2\mu}{K^2}\int\f{\diff^3 k}{(2\pi)^3}\abs{\f{2\pi r_X^2 p_0}{k\cos\theta}\sin\p{\f{h}{2}k\cos\theta}}^2\,,\\
&=\ps{\f{1}{2h}(r_X^2p_0)^2\f{\lambda+2\mu}{K^2}}\ps{hk_0\cos(hk_0)+\sin(hk_0)+hk_0\p{-2+hk_0\te{Si}(hk_0)}}\,,\\
&=\ps{(r_X^2p_0)^2\f{\lambda+2\mu}{K^2}}\ps{\f{\pi}{\lambda_0}\cos\p{\f{2\pi h}{\lambda_0}}+\sin\p{\f{2\pi h}{\lambda_0}}+\p{\f{\pi}{\lambda_0}}\p{-2+\p{\f{2\pi h}{\lambda_0}}\te{Si}\p{\f{2\pi h}{\lambda_0}}}}\,,\\
&\approx_{h\gg\lambda_0\gg1}\ps{(r_X^2p_0)^2\f{\lambda+2\mu}{K^2}}\ps{\f{\pi^3 h}{\lambda_0^2}}\,,\\
&=\ps{(\sigma_Xp_0)^2\f{\lambda+2\mu}{K^2}}\ps{\f{\pi h}{\lambda_0^2}}\,.
\end{align}
From this we obtain the fractional energy deposition into the unattenuated wavelengths
\begin{align}
\Xi&=\f{\sigma_X}{4\pi\lambda_0^2}\,.
\end{align}
This approximation holds for $\lambda_0^2\gg\sigma_X$, which is appropriate for the case that $\lambda_0$ is on the order of kilometers and $\sigma_X$ is on the order of centimeters squared.
\end{document}
